Lake Photic Zone Temperature accross the Conterminous United States

### 2019-08-26 manuscript/manuscript.Rmd
Added Stephen and Bryan's edits

Using the 2007 and 2012 NLA data, we have successfully built a simple yet robust model of lake surface temperature for the conterminous United States. The final model had a mean square error of 2.19 and an adjusted R^2 of 0.88.  The sampling date and that day’s average ambient air temperature, data obtained from the PRISM Climate Group, were the most important variables impacting the final model’s accuracy.  Assuming that ambient air temperature date is available, this model allows us to predict lake photic zone temperature (Importance of this ability??).

@@

An advantage of this model is that we are now able to forecast and backcast lake photic zone temperature.  While there are numerous lakes that have substantial historical temperature data, these data are not consistently available for large regions.  This inhibits our ability to make large scale predictions about changes to lake dynamics (e.g., stratification and mixing) and ecological risk (e.g., anticipated changes in cyanobacteria bloom frequency).  At the continental scale, essentially, we could identify area that might have a higher ecological risk.  Yet if there was an individual lake of concern that lack temperature, this model would be as effective.

Final paragraph

>>>>

Using the 2007 and 2012 NLA data, we have successfully built a simple yet robust model of lake surface temperature for the conterminous United States. The final model has a mean square error of 2.19 and an adjusted R^2 of 0.88. The sampling date, that day’s average ambient air temperature, data obtained from the PRISM Climate Group, and longitude are the most important variables impacting the final model’s accuracy. Given the importance of temperature to a lake ecosystem, especially cyanobacteria bloom dynamics, this model can be a valuable tool for researchers and lake resource managers.  Daily predicted lake photic zone temperature for all lakes in the conterminous US can now be straightforwardly estimated based on basic ambient temperature and location information.

@@

Despite being one of the most common measurements collected by limnologists, lake temperature datasets that cover long periods of time are very difficult to obtain.  Sharma et al (2015) have compiled summer lake temperature data for 291 lakes for the period 1985-2009.  This may be the largest lake temperature database to date, however, the data are only available to members of their research group and, realistically, the number of lakes included is very small.  One of the reasons we chose to model lake photic zone temperature was to develop a database of lake temperatures for the 48 conterminous United States.  The model we present has proven to be accurate and will allow us to backcast lake temperatures for all the > 300,000 lakes included in NHDplus for the period of time covered by the PRISM climate predictions (1981 to present).  This dataset will allow us to investigate how photic zone temperatures vary both spatially and temporally across the United States.  This database is being developed and, when complete, will be made available as an open source data set.
